
---
output: 
  pdf_document:
    fig_caption: false
header-includes:
- \usepackage[utf8]{inputenc}
- \usepackage[T1]{fontenc}
- \usepackage{textcomp}
- \usepackage[english]{babel}
- \usepackage{dcolumn}
- \usepackage{enumitem}
- \usepackage{xcolor}
# geometry: top=0.5cm,left=2cm, right = 2cm
fontsize: 11pt
params:
    name: Migla Miskinyte #Change your name!!
    date: 2021-03-15 #Change date!!
    description: Survival analysis #Change to description of analysis!!
    data: folders/HMI/Migla/data/2021/March #change it 
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = T,background = '#ff9900',highlight = TRUE,comment = NA,warning = F)
```

<!-- DO NOT CHANGE: JUST CHANGE graphics path -->
\begin{minipage}[t]{0.2\textwidth}
\graphicspath{{C:/Migla/}}    
\centering\includegraphics[width=10in]{Wolbachia.png}
\end{minipage}
\begin{minipage}[t]{0.7\textwidth}
\vspace{-2.7cm}
\centering \Large \textbf{Host-microbe interactions 2020/21}\\Date: `r params$date`\\Data in: `r params$data`\\`r params$name`\\
\end{minipage}
<!-- END -->

<!-- DO NOT CHANGE -->
\textbf{DCV pricking test - `r params$description`} 
<!-- END OF DO NOT CHANGE -->

This is my first trial for DCV pricking using two different genotypes of Wolbachia (Iso and wMel) and different concentrations of virus.

\ First, all necessary packages are uploaded. Remove hashtag before running. 

```{r}
#library(tidyverse)
#library(coxme)
#library(car)
#library(emmeans)
#library(here)
#library(ggplot2)

```